The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs

### 2021-09-03 src/paper/report.tex
Changes requested by manuscript checker.

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many months---and
often years---away from reaching vaccination levels that would end the pandemic or even
protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is of utmost
importance to employ an effective mix of strategies for containing the virus. The most
frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
contacts between individuals. While this has allowed some countries to sustain equilibria
with very low infection numbers,\footnote{See \citet{Contreras2021} for a theoretical
equilibrium at low case numbers which is sustained with test-trace-and-isolate policies.}
most have seen large fluctuations of infection rates over time. Containment measures have
become increasingly diverse and now include rapid testing, more nuanced NPIs, and contact
tracing. Neither these policies' effects nor the influence of seasonal patterns or of
more infectious virus strains are well understood in quantitative terms.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests
yield immediate results. Specificity and
sensitivity of these tests is set
according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the
timing of the test relative to the onset of infectiousness. After a phase-in period, all
tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests.\footnote{Positive test results or symptoms lead most individuals to reduce their
    contacts; this is why tests impact the actual contacts in
    Figure~\ref{fig:model_description}.}

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and calibrate the model's many
parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
different contact types and their assortativity by age group. Mobility data is used to
model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}. In our
application, we keep the number of free parameters low in order to avoid overfitting.
The data features to be matched include official case numbers for each age group and
region, deaths, and the share of the B.1.1.7 strain.

@@

The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value one represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of ten. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed, so were customer-facing businesses except
for grocery and drug stores. From the peak of the second wave just before Christmas
until the trough in mid-February, newly detected cases decreased by almost three
quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
(Delta) was first detected in Germany in April; at the end of our simulation period it
accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
overall stringency index.

@@

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in only in mid-March, they make up for more than half of the total effect.
The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not have been nearly as effective.

>>>>

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination levels that would end the
pandemic or even protect the most vulnerable \citep{Mathieu2021}. In the meantime, it is
of utmost importance to employ an effective mix of strategies for containing the virus.
The most frequent initial response was a set of non-pharmaceutical interventions (NPIs)
to reduce contacts between individuals. While this has allowed some countries to sustain
equilibria with very low infection numbers (see \citet{Contreras2021} for a theoretical
equilibrium at low case numbers which is sustained with test-trace-and-isolate policies)
most have seen large fluctuations of infection rates over time. Containment measures
have become increasingly diverse and now include rapid testing, more nuanced NPIs, and
contact tracing. Neither these policies' effects nor the influence of seasonal patterns
or of more infectious virus strains are well understood in quantitative terms.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
sensitivity depends on the timing of the test relative to the onset of infectiousness.
After a phase-in period, all tests that are demanded will be performed.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require staff and students to be tested regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid tests
(positive test results or symptoms lead most individuals to reduce their contacts; this
is why tests impact the actual contacts in Figure~\ref{fig:model_description}).

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and calibrate the model's many
parameters (see Supplementary Material~\ref{sec:materials_and_methods} for a complete
description). Contact diaries yield pre-pandemic distributions of contacts for different
contact types and their assortativity by age group. Mobility data is used to model the
evolution of work contacts. School and daycare policies can be incorporated directly
from official directives. Administrative records on the number of tests, vaccinations by
age and region, and the prevalence of virus strains are generally available. Surveys may
ask about test offers, propensities to take them up, and past tests. Other studies'
estimates of the seasonality of infections can be incorporated directly. The remaining
parameters---most notably, these include infection probabilities by contact network and
the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}. In our
application, we keep the number of free parameters low in order to avoid overfitting.
The data features to be matched include official case numbers for each age group and
region, deaths, and the share of the B.1.1.7 strain.

@@

The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value one represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of ten. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed, so were customer-facing businesses except
for grocery and drug stores. From the peak of the second wave just before Christmas
until the trough in mid-February, newly detected cases decreased by almost three
quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
strain, which became dominant in March (Figure~\ref{fig:share_b117}, note that B.1.617.2
(Delta) was first detected in Germany in April; at the end of our simulation period it
accounted for less than 5\% of cases.). In early March, some NPIs were relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
overall stringency index.

@@

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students (18\% of our population are in the education sector (pupils,
teachers, etc.); 46\% are workers outside the education sector). Almost 40\% come from
tests at the workplace. Despite the fact that rapid tests for private reasons are phased
in only in mid-March, they make up for more than half of the total effect. The reason
lies in the fact that a substantial share of these tests is driven by an elevated
probability to carry the virus, i.e., showing symptoms of CoViD-19 or following up on a
positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not have been nearly as effective.
